---
title: "First small steps with sparklyr on the Kulturarvscluster"
author: Per Møldrup-Dalum (PMD)
date: 13 September 2017
output: html_document
---

## Introduction

To follow this tutorial you need to have the Github project cloned or extracted on the DeIC National Cultural Heritage Cluster, Royal Danish Library[^1] (Kulturarvscluster).

A minimal example on how to get startet with Sparklyr. Focus will be on using R and Spark on data from Netarkivet-ETL.

If you have not yet installed Sparklyr, you need to go through the `first-sparklyr.Rmd` tutorial.

You can then load the Sparklye library

```{r}
library(sparklyr)
```

We also need the functions from the dplyr library.

```{r}
library(dplyr)
```

## Connect to the Spark cluster

When using sparklyr, you can either connect to a local instance of Spark or connect to a remote Spark on a real cluster. In these examples we will focus on submitting jobs to the Kulturarvscluster Spark.

Sparklyr includes a function to install a local Spark. We choose to install Spark in version 2.1.0 using this simple function call

```{r}
spark_install(version = "2.1.0")
```

## Test the sparklyr

Let us try to load some data into Spark and then perform a small query. Just to make it a bit interesting, I choose to use the same data set as on the RStudio introduction to sparklyr[^1]. R comnes with quite a large set of data sets, but we'll work with the New York City Flights 2013 dataset which needs to be installed

```{r}
install.packages("nycflights13")
library(nycflights13)
flights
```

This data set contains 336,776 flights out of New York in 2013.

We need just one more component for this test. If you once have tasted the sweet taste of dplyr, you can never go back——which is not that bad. Alas,

```{r}
library(dplyr)
```

Okay, now it's time to create a Spark context

```{r}
sc <- spark_connect(master = "local")
```
Put the flights into Spark

```{r}
flights_tbl <- copy_to(sc, flights, "flights", overwrite = TRUE)
src_tbls(sc)
```

Okay, first let's try a simple query. Let's say we want to look at flights with a departure delayed two hours and group those by day

```{r}
flights_tbl %>%
    filter(dep_delay == 2) %>%
    group_by(month) %>%
    summarise(n = n()) %>%
    arrange(month)
```

Now that we at it, we might as do the same plot that RStudio uses to demonstrate sparklyr

```{r}
delay <- flights_tbl %>%
    group_by(tailnum) %>%
    summarise(count = n(), dist = mean(distance), delay = mean(arr_delay)) %>%
    filter(count > 20, dist < 2000, !is.na(delay)) %>%
    collect

library(ggplot2)
ggplot(delay, aes(dist, delay)) +
    geom_point(aes(size = count), alpha = 1/2) +
    geom_smooth() +
    scale_size_area(max_size = 2)
```
And that's it, Folks...

[^1]: https://spark.rstudio.com


[^1]: http://kulturarvscluster.dk
